\Huge{Vipassana for Hackers} \\ \large{Paper Two: The Brain} \\ \large\textit{Version 0.2}

### 2019-09-29 paper-two/paper-two.tex

The DMN is comprised of specific anatomy including portions of the mid-line of the
brain, an evolutionarily primitive area related to memory and emotion, and
structures in the cortex, an evolutionarily recent part of the brain containing the
executive and higher order functions. \cite{defaultnetworkanatomy} These functionally
connected regions are involved in the neurological basis
of the self, considering the mental states of others, remembering the past, and
imagining the future.

@@

The regions involved in an individual's conceptual ``self'' are the Posterior
Cingulate Cortex (PCC), Ventromedial Prefrontal Cortex (vmPFC), and the Inferior
Parietal Lobule (IPL). \cite{defaultnetworkadaptive}

Cerebral blood flow and metabolic rate is 40\%
higher in the PCC compared to the average across the brain, making it one of the most
active regions of the brain. \cite{pccrole}

The PCC has been thought to mediate interactions between emotion and memory. Under
fMRI, it consistently lights up when a person recollects something from their
life. The strength of PCC brain activity in such an experiment varies depending on
the emotional importance of the memory recalled. It is also activated by emotional
stimuli, both positive and negative, and acts as an interface between the external
world and the individual by gauging the importance of different
stimuli. \cite{pccemotion}

@@

The vmPFC is involved in creating a conceptual self by self-related processing
and the assignment of personal significance to self-related
information. \cite{dmnself} The vmPFC is paired with the Dorsomedial Prefrontal
Cortex (dmPFC), which assists in the creation of the conceptual self through an
individual's consideration of ``others'', to construct the Medial Prefrontal Cortex
(mPFC).

\todo{explain IPL?}

It is hypothesized that the activation of these three regions of the brain (PCC,
vmPFC, and IPL) are responsible for providing a sense of self --- a subjective agent
in space and time. It functions as a network in which phenomena such as self
representations are accessible via the PCC \todo{add PCC reference} and are gated into conscious awareness by
activity in the mPFC, as influenced by changing internal and external demands. The
PCC acts as a brain-wide connectivity hub, through which a unitary notion of self is
created by a large scale integration of DMN activity. \cite{mappingself}

@@

>>>>

The DMN is comprised of specific anatomy including portions of the mid-line of the
brain, an evolutionarily primitive area related to memory and emotion, and
structures in the cortex, an evolutionarily recent part of the brain containing the
executive and higher order functions. \cite{defaultnetworkanatomy} These functionally
connected regions are involved in the neurological basis
of the self, considering the mental states of others, remembering the past, and
imagining the future. 
This network is mostly observed through changes in blood flow to different parts of 
the brain measured using fMRI (functional Magnetic Resonance Imaging) or 
PET (Positron Emission Tomography){defaultnetworkadaptive}

@@

The regions involved in an individual's conceptual` `self'' are the Posterior
Cingulate Cortex (PCC), Ventromedial Prefrontal Cortex (vmPFC), and the Inferior
Parietal Lobule (IPL).That is to say, these regions are preferentially activated when 
volunteers were asked to reflect on personal preferences, beliefs, abilities, 
anything relevant to the self. \cite{defaultnetworkadaptive}

Cerebral blood flow and metabolic rate is 40\% higher than average in the PCC,
making it one of the most active regions of the brain. \cite{pccrole}

The PCC has been thought to mediate interactions between emotion and memory. 
Under an fMRI, it consistently lights up when a person recollects something from their
life. The strength of this activity varies depending on the emotional importance of 
the memory recalled. It is also activated by emotional stimuli, 
both positive and negative, and acts as an interface between the external world 
and the individual by gauging the importance of different stimuli. \cite{pccemotion}

@@

The vmPFC is involved in creating a conceptual self by self-related processing
and the assignment of personal significance to self-related
information. \cite{dmnself} The vmPFC is paired with the Dorsomedial Prefrontal
Cortex (dmPFC), which assists in the creation of the conceptual self through an
individual's consideration of ``others''.

The Inferior Parietal lobe (IPL) is situated in the back of the brain and it is 
involved in the perception of facial stimuli.\cite{IPL}

It is hypothesized that the activation of these three regions of the brain (PCC,
vmPFC, and IPL) are responsible for providing a sense of self --- a subjective agent
in space and time. It functions as a network in which phenomena such as self
representations are accessible via the PCC \cite{mappingself} and are gated into 
conscious awareness by activity in the mPFC, as influenced by changing internal 
and external demands. The PCC acts as a brain-wide connectivity hub, through which a unitary notion of self is
created by a large scale integration of DMN activity. \cite{mappingself}

@@

\bibitem{IPL}
 Radua J, Phillips ML, Russell T, Lawrence N, Marshall N, Kalidindi S, El-Hage W, McDonald C, 
 Giampietro V, Brammer MJ, David AS, Surguladze SA.
 \textit{Neural response to specific components of fearful faces in healthy 
  and schizophrenic adults.}
  Neuroimage. 2010 Jan 1;49(1):939-46. doi: 10.1016/j.neuroimage.2009.08.030. 
  Epub 2009 Aug 20.
  \url:https://www.ncbi.nlm.nih.gov/pubmed/19699306
